A molecular dynamics study of the solvation of \ce{CO_2}, water, and some organic compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-07-23 ionic_liquids.tex
intro small corrections

\begin{abstract}
In this paper we calculate solvation free enegies in ionic liquids and predict macroscopic properties required in the design of chemical processes. Carbon dioxide produced in combustion processes is one of the main sources of anthropogenic generation of greenhouse gases. On the other hand, the potential of \ce{[emim][B(CN)_4]} for its capture has recently been reported. For comparison purposes, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, namely the pressure required to obtain the same molality of \ce{CO_2} is notably lower in \ce{[emim][B(CN)_4]}. The infinite-dilution activity coefficients of hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.
\end{abstract}

@@

\section{Introduction}

The \ce{CO_2} emissions from power generation plants are still an important source of anthropogenic greenhouse gases \cite{totalenergy}. International efforts aimed at regulating and reducing \ce{CO_2} emissions are increasing \cite{Tong_2018} and, in this context, the development of technologies for the capture and storage of \ce{CO_2} is essential. \cite{Markewitz_2014}.

The conventional process for \ce{CO_2} removal involves a chemical absorption followed by stripping at high temperatures. The solvents commonly employed are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA). Despite its high absorption capability, the use of MEA suffers from a serious drawback, namely a considerable cost due to the energy consumption of solvent recovery \cite{Merkel_2010}. Other disadvantages include corrosion and degradation by oxidation, solvent loss by evaporation, which also contributes to increase the ambient pollution and generates extra costs to replace the solvent.

The use of ionic liquids (ILs) as solvents for \ce{CO_2} sequestration occupies a central place in the investigation of alternative technologies. The reasons are their well known properties, such as the negligible vapor pressure and high thermal stability. Furthermore, they are particularly attractive because they can be designed for specific applications, giving rise to the so-called task-specific ILs \cite{Seo_2014}. These remarkable properties makes ILs suitable for a wide variety of applications in separation processes \cite{Han_2010,Werner_2010}, catalysis \cite{P_rvulescu_2007}, energy \cite{MacFarlane_2014} and material science \cite{Mecerreyes_2011,Tom_2015,Dupont_2010,Leones_2017,Kinik_2017}.

Recently, the potentiality of ILs for the \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs. While most thermodynamic models were developed based on molar fraction, the key factor to determine the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain volume or mass of solvent. By expressing the solubility in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a \ce{CO_2} solubility similar or even lower than that observed for the heavy alkane eicosane. This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}, the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-]}.

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of the anion \ce{[B(CN)_4]^-]}, whether special interactions with \ce{CO_2} are settled. In this sense, Molecular Simulation provides with a powerful tool for studying solvation processes. To the best of our knowledge, only one study in the literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated. This motivated us to carry out a comparison of different force fields of \ce{[emim][B(CN)_4]} in terms of the free energy of solvation and, indeed,  it was possible to select a model that successfully predicts the value of the Henry constant of \ce{CO_2}. For comparison, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, that is, a considerably lower pressure is required in \ce{[emim][B(CN)_4]} to obtain the same molality of \ce{CO_2}. The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in the literature.

This paper is organized as follows: in Sec. \ref{sec:theory} we present the mathematical relationships between the simulated solvation free energy and thermodynamic properties. This is followed by a brief description of the force fields considered for the ILs and the solutes. The methodological details are provided in Sec. \ref{sec:sim_details} and the numerical results are given in Sec. \ref{sec:results}. Finally, we present some concluding remarks in Sec. \ref{sec:conclusion}.

@@

In this section, we present the formalism that allows a direct comparison between simulation results and the thermodynamic properties of dilute solutions: Henry constant and activity coefficient at infinite dilution. For this, we rely on the work of Shirts \textit{et al.} \cite{Shirts_2003}. Then, we reformulate the equations in order to carry out a study at concentrated conditions, which would be the more likely scenario in \ce{CO_2} capture processes.

>>>>

\begin{abstract}
In this paper, we calculate solvation free enegies in ionic liquids and predict macroscopic properties required in the design of chemical processes. Carbon dioxide produced in combustion processes is one of the main sources of anthropogenic generation of greenhouse gases. On the other hand, the potential of \ce{[emim][B(CN)_4]} for its capture has recently been reported. For comparison purposes, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, namely the pressure required to obtain the same molality of \ce{CO_2} is notably lower in \ce{[emim][B(CN)_4]}. The infinite-dilution activity coefficients of hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.
\end{abstract}

@@

\section{Introduction}
\ce{CO_2} emissions from fossil-fuel power generation plants account for a large portion of the anthropogenic greenhouse gases \cite{totalenergy}. International efforts aimed at regulating and reducing \ce{CO_2} emissions are increasing \cite{Tong_2018} and, in this context, the development of technologies for the capture and storage of \ce{CO_2} plays an essential role \cite{Markewitz_2014}.

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures. The most commonly used solvents are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA). Despite its high absorption capability, the use of MEA suffers from a serious drawback due to the high cost of solvent recovery \cite{Merkel_2010}. Other disadvantages include corrosion and degradation by oxidation, solvent loss by evaporation, which also contributes to increase the environment pollution and generates extra costs for replacing the solvent.

The use of ionic liquids (ILs) as solvents for \ce{CO_2} sequestration occupies a central place in the investigation of alternative technologies. The reasons are their well known properties, such as their almost negligible vapor pressure and high thermal stability. Furthermore, they are particularly attractive since they can be designed for specific applications, giving rise to the so-called task-specific ILs \cite{Seo_2014}. These remarkable properties makes ILs suitable for a wide variety of applications in separation processes \cite{Han_2010,Werner_2010}, catalysis \cite{P_rvulescu_2007}, energy \cite{MacFarlane_2014}, and material science \cite{Mecerreyes_2011,Tom_2015,Dupont_2010,Leones_2017,Kinik_2017}.

Recently, the potentiality of ILs for \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs. Although most thermodynamic models are formulated in terms of molar fractions, a key factor that dictates the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent. By expressing the solubility values in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a solubility similar or even lower than that observed for the heavy alkane eicosane. This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}; the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-}.

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of the anion \ce{[B(CN)_4]^-}. In this sense, Molecular Simulation provides with a powerful tool for studying solvation processes. To the best of our knowledge, only one study in literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated. This motivated us to evaluate different force fields of \ce{[emim][B(CN)_4]} and, as we will show below, it was possible to select a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2}. For comparison, \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what is observed in experiments: for \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}. The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.

This paper is organized as follows: in Sec.~\ref{sec:theory} we present the theoretical background, which is followed by a brief description of the force fields used in this work. The methodological details are provided in Sec.~\ref{sec:sim_details} and the corresponding numerical results are given in Sec.~\ref{sec:results}. Finally, we present some concluding remarks in Sec.~\ref{sec:conclusion}.

@@

In this section, we present the formalism that allows us to compute the Henry constant and activity coefficient, both at infinite dilution, from the simulated free energy. For this, we rely on the work of Shirts \textit{et al.} \cite{Shirts_2003}. Then, we reformulate the equations in order to study concentrated solutions, which would be the more likely scenario in \ce{CO_2} capture processes.
